Numerical Solution of Point Defect Equations by MOOSE

### 2021-01-19 examples/point_defects/article/article.tex
article-edit

\documentclass[a4paper]{article}
\usepackage[margin=25mm]{geometry}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage[font=scriptsize]{caption}
\pagenumbering{arabic}
\graphicspath{ {./plots/} }

@@

@@

@@

    The vacancy supersaturation curves for biased neutron irradation is given in Fig. \ref{figure:vacancy_supersaturation_neutron_5}. Since the defect concentrations have symmetrical profiles for neutron irradation, supersaturation curves are also symmetrical around grain center. The plot shows a non-linear increase in supersaturation values as grain gets larger.
      \begin{figure}[h!]  %Nuetron - Vacancy Supersaturation - 5% Bias
        \centering
        \subfloat{a)\includegraphics[scale=0.55]{super_saturation_500-5000nm-neutron-5}}
        \qquad
        \subfloat{b)\includegraphics[scale=0.55]{super_saturation_500-5000nm-high_neutron-5}}
        \caption{Vacancy supersaturation profiles for 5\% production bias, a)low dose rate neutron, b) high dose rate neutron}
        \label{figure:vacancy_supersaturation_neutron_5}
      \end{figure}

    \newpage
    \subsubsection{Sink Strengths} \hspace{10pt}
    The sink reaction term in Eq. \ref{equation:point_defect_equations_sink_strength} represents the total loss of point defects to sinks that are voids and dislocations within the domain. It can be written in expanded form as below,

@@

@@

    The vacancy supersaturation curves for ion irradation case are quite similar to ones for neutron irradation case. The only difference is the asymmetrical shape of curves originates from concentration profiles. The non-linear increase in supersaturation values as domain gets larger is observed for ion irradation too.
      \begin{figure}[h!]  %Ion - %5 Bias rate
        \centering
        \subfloat{a)\includegraphics[scale=0.55]{super_saturation_500-5000nm-low_ion-5}}
        \qquad
        \subfloat{b)\includegraphics[scale=0.55]{super_saturation_500-5000nm-ion-5}}
        \caption{Vacancy supersaturation profiles for 5\% production bias, a) low ion dose rate, b) high ion dose rate}
        \label{figure:vacancy_supersaturation_ion_5}
      \end{figure}
      \newpage
    \subsubsection{Sink Strengths} \hspace{10pt}
    The sink strengths are directly related to concentration at center and flow at boundary. The change in centerline defect concentrations are not as considerable as neutron case. So, the flow or flux at boundary becomes more dominant on sink strength profiles (see Fig. \ref{figure:sink_strengths_ion_5_1e-6} and \ref{figure:sink_strengths_ion_5_1e-3}).
      % \begin{figure}[h!]  %Sink Strength - ion 1e-3 dpa/s - 0% Bias rate
      %   \centering
      %   \subfloat{\includegraphics[scale=0.55]{sink_strength_moose_ion_0_k}}
      %   \qquad
      %   \subfloat{\includegraphics[scale=0.55]{sink_strength_moose_ion_0_Z}}
      %   \caption{Sink Strengths for low dose ion irradation with 0\% Production Bias}
      %   \label{figure:sink_strengths_ion_0_1e-6}
      % \end{figure}
      \begin{figure}[h!]  %Sink Strength - ion 1e-6 dpa/s - 5% Bias rate
        \centering
        \subfloat{\includegraphics[scale=0.55]{sink_strength_moose_low_ion_5_Z_scaled}}
        \qquad
        \subfloat{a)\includegraphics[scale=0.55]{sink_strength_moose_ion_5_Z_scaled}}
        \caption{Change of GB sink strength with grain size for 5\% production bias, a) low ion dose rate, b) high ion dose rate}
        \label{figure:sink_strengths_ion_5_1e-6}
      \end{figure}
      \begin{figure}[h!]  %Sink Strength - ion 1e-3 dpa/s - 5% Bias rate
        \centering
        \subfloat{\includegraphics[scale=0.55]{sink_strength_moose_low_ion_5_k_scaled}}
        \qquad
        \subfloat{b)\includegraphics[scale=0.55]{sink_strength_moose_ion_5_k_scaled}}
        \caption{Change of total GB sink strength with grain size for 5\% production bias, a) low dose rate neutron, b) high dose rate neutron}
        \label{figure:sink_strengths_ion_5_1e-3}
      \end{figure}

% In nano-scale case, since the grain size is not big enough, the boundaries become
% closer to each other. This causes point defects to reach sinks before they react with each other
% and recombination plateau to disappear as it can be seen in Fig. 2. The nanoscale case acts like
% a high-sink and high-temp condition case despite its low-sink and low-temperature. Therefore,
% change in production bias does not have a significant effect on average concentration behavior
% in nanoscale
%
% On the other hand, in micro-scale case, the domain has reasonable size, and thus
% boundaries are far from each other. This provides defects more time to accumulate inside the
% domain and recombine with each other after they reach a specific concentration. Figure 3
% clearly shows the effect of production bias on average concentration.
%
% Effect of production bias on spatial profile of defect concentrations for nanoscale is
% shown in Figure 4. The results seem not affected too much because of the same reasons
% mentioned previously for average concentration.
%
% The micro-scale results (Fig. 5) are quite interesting and worth to think over. The
% concentration profiles of interstitials for 0.1%, 5% and 10% cases are flipped over. A closer look
% at evolution of 5% case over time in Fig. 6, after the end of quasi-steady-state region (t=10 -4 s in
% Fig 3.), the average interstitial concentration starts to decrease as they segregate through sinks.
% At steady state, the average interstitial concentration becomes minimum at the center and
% maximum near boundaries because of segregation.
%
% The defect production rate (K 0 ) is the source term in point defect equations and has a
% significant effect on concentrations.
% In nano-scale case, the average concentration behavior is like high-sink density behavior
% because of narrow domain. But, when the defect production rate increases, the average
% concentration is shifted upward, and interstitial and vacancy average concentrations get closer
% to each other because the production compensates more absorption at sinks than previous.
% Based on this, if the production rate increased sufficiently, the recombination zone would
% appear again in nano-scale domain.
  \subsection{Discussion} \hspace{10pt}
  What happens when switched from neutron irradation to ion irradiation
  \begin{enumerate}
    \item Loosing symmetry/uniformity in concentration profiles
    \item Reduction in average concentration
    \item Reduction in supersaturation
  \end{enumerate}

\newpage
\section{Conclusion} \hspace{10pt}

@@

>>>>

\documentclass[a4paper]{article}
\usepackage[margin=25mm]{geometry}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{xcolor}
\usepackage[font=scriptsize]{caption}
\pagenumbering{arabic}
\graphicspath{ {./plots/} }

@@

\textcolor{red}{!!!!!!!!!!!!!!!!!!!!!!!ADD PARAGRAPH!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!}

@@

    \begin{figure}[htb!]  %Neutron - 1e-6 dpa/s - %5 Bias rate
      \centering
      \subfloat{a)\includegraphics[scale=0.28]{data_neutron_250nm}}
      \qquad
      \subfloat{b)\includegraphics[scale=0.28]{data_neutron_1000nm}}
      \caption{Change of interstitial concentration profiles with production bias for low dose rate neutron in a grain of a) 250 $nm$ and b) 1000 $nm$ }
      \label{figure:3D_concentrations_neutron_1e-6}
    \end{figure}

@@

    The vacancy supersaturation curves for biased neutron irradation is given in Fig. \ref{figure:vacancy_supersaturation_neutron_5}. Since the defect concentrations have symmetrical profiles for neutron irradation, supersaturation curves are also symmetrical around grain center. The plot shows a non-linear increase in supersaturation values as grain gets larger.
      \begin{figure}[h!]  %Neutron - Vacancy Supersaturation - 5% Bias
        \centering
        \subfloat{a)\includegraphics[scale=0.55]{super_saturation_500-5000nm-neutron-5}}
        \qquad
        \subfloat{b)\includegraphics[scale=0.55]{super_saturation_500-5000nm-high_neutron-5}}
        \caption{Vacancy supersaturation profiles for 5\% production bias, a)low dose rate neutron, b) high dose rate neutron}
        \label{figure:vacancy_supersaturation_neutron_5}
      \end{figure}

      \begin{figure}[htb!]  %Neutron - Vacancy Supersaturation
        \centering
        \subfloat{a)\includegraphics[scale=0.38]{data_neutron_Sv_center}}
        \qquad
        \subfloat{b)\includegraphics[scale=0.38]{data_high_neutron_Sv_center}}
        \caption{Change of vacancy supersaturation at center with grain size and production bias for a)low dose rate neutron, b) high dose rate neutron}
        \label{figure:3D_supersaturation_neutron}
      \end{figure}
    \newpage
    \subsubsection{Sink Strengths} \hspace{10pt}
    The sink reaction term in Eq. \ref{equation:point_defect_equations_sink_strength} represents the total loss of point defects to sinks that are voids and dislocations within the domain. It can be written in expanded form as below,

@@

      \begin{figure}[htb!]  %Neutron - 1e-6 dpa/s - %5 Bias rate
        \centering
        \subfloat{a)\includegraphics[scale=0.28]{data_low_ion_250nm}}
        \qquad
        \subfloat{b)\includegraphics[scale=0.28]{data_low_ion_1000nm}}
        \caption{Change of interstitial concentration profiles with production bias for low dose rate ion in a grain of a) 250 $nm$ and b) 1000 $nm$ }
        \label{figure:3D_concentrations_ion_1e-6}
      \end{figure}

@@

    The vacancy supersaturation curves for ion irradation case are quite similar to ones for neutron irradation case. The only difference is the asymmetrical shape of curves originates from concentration profiles. The non-linear increase in supersaturation values as domain gets larger is observed for ion irradation too.
      \begin{figure}[h!]  %Ion - %5 Bias rate
        \centering
        \subfloat{a)\includegraphics[scale=0.55]{super_saturation_500-5000nm-low_ion-5}}
        \qquad
        \subfloat{b)\includegraphics[scale=0.55]{super_saturation_500-5000nm-ion-5}}
        \caption{Vacancy supersaturation profiles for 5\% production bias, a) low ion dose rate, b) high ion dose rate}
        \label{figure:vacancy_supersaturation_ion_5}
      \end{figure}

      \begin{figure}[htb!]  %Neutron - Vacancy Supersaturation
        \centering
        \subfloat{a)\includegraphics[scale=0.38]{data_low_ion_Sv_center}}
        \qquad
        \subfloat{b)\includegraphics[scale=0.38]{data_ion_Sv_center}}
        \caption{Change of vacancy supersaturation at center with grain size and production bias for a) low dose rate ion, b) high dose rate ion}
        \label{figure:3D_supersaturation_ion}
      \end{figure}

      The vacancy supersaturation behavior, shown in Fig. \ref{figure:3D_supersaturation_ion}, is identical to one in given for neutron irradiation case (see Fig. \ref{figure:3D_supersaturation_neutron}). This similarity makes it possible to use ions instead of neutrons to analyze swelling of reactor materials.

      \newpage
    \subsubsection{Sink Strengths} \hspace{10pt}
    The sink strengths are directly related to concentration at center and flow at boundary. The change in centerline defect concentrations are not as considerable as neutron case. So, the flow or flux at boundary becomes more dominant on sink strength profiles (see Fig. \ref{figure:sink_strengths_ion_5_1e-6} and \ref{figure:sink_strengths_ion_5_1e-3}).
      \begin{figure}[h!]  %
        \centering
        \subfloat{a)\includegraphics[scale=0.55]{interstitial_concentration_5-5000nm-low_ion-5}}
        \qquad
        \subfloat{b)\includegraphics[scale=0.55]{interstitial_concentration_5-5000nm-ion-5}}
        \caption{Interstitial concentration profiles with 5\% production bias for a) low ion dose rate, b) high ion dose rate}
        \label{figure:concentrations_ion_5_1e-3_5-5000nm}
      \end{figure}

      As it is discussed previously, since ions can not penetrate till the end of domain, they cause a non-uniform defect production. If the grain is small enough compared to penetration depth, concentration profiles are symmetrical, and the effect of distribution profile is not visible (up to 1000 $nm$ in Fig. \ref{figure:concentrations_ion_5_1e-3_5-5000nm}a and 50 $nm$ in Fig. \ref{figure:concentrations_ion_5_1e-3_5-5000nm}b). Yet, for larger grains, the defect concentrations are highly affected by non-uniformity. The flux at the far boundary oscillates and approaches to zero as grain size increases. Since sink strengths are related to the ratio of flux to concentration, they are affected by the oscillations in flux and concentration (see Fig. \ref{figure:sink_strengths_ion_5_1e-6}).

      \begin{figure}[h!]  %Sink Strength - ion 1e-6 dpa/s - 5% Bias rate
        \centering
        \subfloat{a)\includegraphics[scale=0.55]{sink_strength_moose_low_ion_5_Z_scaled}}
        \qquad
        \subfloat{b)\includegraphics[scale=0.55]{sink_strength_moose_ion_5_Z_scaled}}
        \caption{Change of GB sink strength with grain size for 5\% production bias, a) low ion dose rate, b) high ion dose rate}
        \label{figure:sink_strengths_ion_5_1e-6}
      \end{figure}
      \begin{figure}[h!]  %Sink Strength - ion 1e-3 dpa/s - 5% Bias rate
        \centering
        \subfloat{a)\includegraphics[scale=0.55]{sink_strength_moose_low_ion_5_k_scaled}}
        \qquad
        \subfloat{b)\includegraphics[scale=0.55]{sink_strength_moose_ion_5_k_scaled}}
        \caption{Change of total GB sink strength with grain size for 5\% production bias, a) low dose rate neutron, b) high dose rate neutron}
        \label{figure:sink_strengths_ion_5_1e-3}
      \end{figure}

      \begin{figure}[h!]  %Sink Strength 3D - Ion 1e-6 dpa/s
        \centering
        \subfloat{a)\includegraphics[scale=0.25]{data_low_ion_Zi}}
        \subfloat{b)\includegraphics[scale=0.25]{data_low_ion_Zv}}
        \qquad
        \subfloat{c)\includegraphics[scale=0.25]{data_low_ion_ki}}
        \subfloat{d)\includegraphics[scale=0.25]{data_low_ion_kv}}
        \caption{Change of sink strengths with grain size and production bias for low dose rate ion in log scale, a)interstitial GB sink strength, b) vacancy GB sink strength, c) total interstitial GB sink strength, d) total vacancy GB sink strength}
        \label{figure:sink_strength_moose_neutron_3D}
      \end{figure}

% In nano-scale case, since the grain size is not big enough, the boundaries become
% closer to each other. This causes point defects to reach sinks before they react with each other
% and recombination plateau to disappear as it can be seen in Fig. 2. The nanoscale case acts like
% a high-sink and high-temp condition case despite its low-sink and low-temperature. Therefore,
% change in production bias does not have a significant effect on average concentration behavior
% in nanoscale
%
% On the other hand, in micro-scale case, the domain has reasonable size, and thus
% boundaries are far from each other. This provides defects more time to accumulate inside the
% domain and recombine with each other after they reach a specific concentration. Figure 3
% clearly shows the effect of production bias on average concentration.
%
% Effect of production bias on spatial profile of defect concentrations for nanoscale is
% shown in Figure 4. The results seem not affected too much because of the same reasons
% mentioned previously for average concentration.
%
% The micro-scale results (Fig. 5) are quite interesting and worth to think over. The
% concentration profiles of interstitials for 0.1%, 5% and 10% cases are flipped over. A closer look
% at evolution of 5% case over time in Fig. 6, after the end of quasi-steady-state region (t=10 -4 s in
% Fig 3.), the average interstitial concentration starts to decrease as they segregate through sinks.
% At steady state, the average interstitial concentration becomes minimum at the center and
% maximum near boundaries because of segregation.
%
% The defect production rate (K 0 ) is the source term in point defect equations and has a
% significant effect on concentrations.
% In nano-scale case, the average concentration behavior is like high-sink density behavior
% because of narrow domain. But, when the defect production rate increases, the average
% concentration is shifted upward, and interstitial and vacancy average concentrations get closer
% to each other because the production compensates more absorption at sinks than previous.
% Based on this, if the production rate increased sufficiently, the recombination zone would
% appear again in nano-scale domain.
\newpage
  \subsection{Discussion} \hspace{10pt}
  What happens when switched from neutron irradation to ion irradiation
  \begin{enumerate}
    \item Loosing symmetry/uniformity in concentration profiles
    \item Reduction in average concentration
    \item Reduction in supersaturation
  \end{enumerate}

\section{Conclusion} \hspace{10pt}

@@

Suplementary materials and extra figures are available at https://github.com/abdurrahmanozturk/articles
